---
title: "DigiGait Analysis"
output: html_document
---

We are only interested in the Right and left hind legs of each mouse and the speeds/angles we care about are the 25cms_0, 25cms_10UP, 25cms_10DN and 30cms_0. I would focus on the %SwingStride column (E). 

Of note: The timepoints between these two groups are different so I've been calling timepoints 8 and 11 the same between the two groups

Typically my pipeline for analysis:

1) Paired T test of R vs L hind legs for each timepoint within each group
2) 1 Way ANOVA of Right hinds overtime for each group
3) T test comparing Right hinds from both groups at each timepoint
occasionally baseline will be significant here and I'll subtract baseline from each latter timepoint, but we can always cross that bridge later
4) The graph I typically make shows just the right hinds from both groups overtime on the same graph for each speed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyr)
library(stringr)
library(broom)
library(ggplot2)
```


```{r}

# list the files in the TLR4_KO folder
TLR4_KO_files <- list.files("../Data/Ariel/TLR4 KO", full.names = TRUE)

# initiate empty data frame
all_gated_df_TLR4 <- data.frame()

for (i in 1:length(TLR4_KO_files)) {
  all_data_TLR4 <- read_xlsx(TLR4_KO_files[i]) %>%
    mutate(Timepoint = TLR4_KO_files[i],
           Group = "TLR4_KO") 
  all_gated_df_TLR4 <- rbind(all_gated_df_TLR4, all_data_TLR4)
}

# list the files in the TLR4_KO folder
WT_files <- list.files("../Data/Ariel/Wild Type", full.names = TRUE)

# initiate empty data frame
all_gated_df_WT <- data.frame()

for (i in 1:length(WT_files)) {
  all_data_WT <- read_xlsx(WT_files[i]) %>%
    mutate(Timepoint = WT_files[i],
           Group = "WT") 
  all_gated_df_WT <- rbind(all_gated_df_WT, all_data_WT)
}

TLR4_WT_all_files <- rbind(all_gated_df_WT, all_gated_df_TLR4)

all_clean <- TLR4_WT_all_files %>%
  filter(Limb == "Right Hind" | Limb == "Left Hind") %>%
  select(FileName, Limb, `%SwingStride`,  Timepoint, Group) %>%
  separate(FileName, into = c("Mouse", "speed", "angle", "hind")) %>%
  select(speed, angle, Mouse, Limb, `%SwingStride`, Timepoint, Group) %>%
  filter(Timepoint != "10.21.15 baseline.xlsx") %>%
  unite(speed_angle, c(speed, angle), sep = "_") %>%
  filter(speed_angle == "25cms_0degUP" | speed_angle == "25cms_10degUP" | speed_angle == "25cms_10degDN" | speed_angle == "30cms_0degUP") %>%
  mutate(Timepoint = str_replace(Timepoint, "../Data/Ariel/TLR4 KO/", ""),
         Timepoint = str_replace(Timepoint, "../Data/Ariel/Wild Type/", "")) %>%
  mutate(Timepoint = str_replace(Timepoint, ".xlsx", "")) %>%
  mutate(`%SwingStride` = as.numeric(`%SwingStride`)) %>%
  mutate(Timepoint = str_replace(Timepoint, "[0-9][0-9]\\.[0-9][0-9]\\.[0-9][0-9]\\s", "")) %>%
  mutate(Timepoint = str_replace(Timepoint, "day 11", "day 8")) %>%
  rename(SwingStride = "%SwingStride") %>%
  mutate(Timepoint = str_replace(Timepoint, "tlr4 day 1 post sx", "day 1")) %>%
  mutate(Timepoint = str_replace(Timepoint, " redo", "")) %>%
  mutate(Timepoint = str_replace(Timepoint, "baseline", "day 0")) %>%
  mutate(Timepoint = str_replace(Timepoint, "day ", "")) %>%
  mutate(Timepoint = as.numeric(Timepoint))

```

Paired T test of R vs L hind legs for each timepoint within each group
```{r}

all_clean %>%
  group_by(Group, Timepoint) %>%
  do(tidy(t.test(.$`SwingStride`~.$Limb, paired = TRUE))) %>%
  filter(p.value < 0.05)

```

1 Way ANOVA of Right hinds overtime for each group
```{r}
all_clean %>%
  filter(Limb == "Right Hind") %>%
  group_by(Group) %>%
  do(tidy(aov(.$`SwingStride`~.$Timepoint))) 
```

T test comparing Right hinds from both groups at each timepoint
occasionally baseline will be significant here and I'll subtract baseline from each latter timepoint, but we can always cross that bridge later.

```{r}

```


The graph I typically make shows just the right hinds from both groups overtime on the same graph for each speed.

```{r}
all_clean_right_hind <- all_clean %>%
  filter(Limb == "Right Hind")  


ggplot(all_clean_right_hind, aes(x = Timepoint, y = SwingStride, color = Group)) +
  geom_point() +
  geom_line() +
  facet_wrap(~speed_angle) +
  theme(axis.text = element_text(angle = 45, hjust = 1))
```

