---
title: "Nanostring Analysis"
author: "Amy Fox"
output: html_document
---

This is a pipeline for analyzing nanostring data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(broom)
library(readr)
library(ggpubr)
```

# Compare LPS v. saline
First need to run normality test on data, if True, then run F.test, if true, then run t.test.

Read in and clean the data
```{r}

data <- read_csv("../Data/Ariel/Normalized GAPDH and Cltc_NormalizedData.csv", skip = 1) %>%
  rename(gene_name= "X2",
         assension_no = "X3") 

data <- data[14:281, ] %>%
  gather(group, count, -Description, -gene_name, -assension_no) %>%
  separate(col = group, into = c("Treatment", "Time", "Sample"), by = " ") %>%
  na.omit() %>%
  mutate(count = as.numeric(count))

```

Normality test: these genes are the ones that are normally distributed
```{r}
normal_test_data <- data %>%
  group_by(gene_name) %>% 
  do(tidy(shapiro.test(.$count)))


normally_distributed <- normal_test_data %>%
  filter(p.value > 0.05)


not_normal_distribution <- normal_test_data %>%
  filter(p.value < 0.05) %>%
  select(gene_name)

# run Mann-Whitney Wilcox on non-normally distributed

if (nrow(not_normal_distribution) > 0) {
data_for_Wilcox <- left_join(not_normal_distribution, data) 

significant_Wilcox <- data_for_Wilcox %>%
  group_by(gene_name) %>% 
  do(tidy(wilcox.test(.$count~.$Treatment))) %>%
  filter(p.value < 0.0500)

wilcox_for_plot <- left_join(significant_Wilcox, data)

ggplot(wilcox_for_plot, aes(Treatment, count)) +
  geom_boxplot(aes(color = gene_name)) +
  facet_wrap(~gene_name, scales = "free") +
  stat_compare_means(method = "wilcox.test", exact = TRUE) +
  ggtitle("Significantly Different Genes (Non-Normal Data)") +
  ylab("Gene Expression") +
  theme(legend.position = "none") }

```

F-test: tests if there are differences in the variances
```{r message = FALSE}
normally_distributed_remove <- normally_distributed %>%
  select(gene_name)

data_for_f_test <- left_join(normally_distributed_remove, data)

f_test_pass <- data_for_f_test %>%
  group_by(gene_name) %>%
  nest() %>%
  mutate(f_test_result = map(data, ~var.test(count ~Treatment, data = .x)),
         tidy_f_test = map(f_test_result, tidy)) %>%
  unnest(tidy_f_test, .drop = TRUE) %>%
  filter(p.value > 0.05)

```
  
T test: 
```{r}
f_test_genes <- f_test_pass %>%
  select(gene_name)

data_for_t_test <- left_join(f_test_genes, data) 

t_test_data <- data_for_t_test %>%
  group_by(gene_name) %>% 
  do(tidy(t.test(.$count~.$Treatment)))

all_passed_data <- t_test_data %>%
  filter(p.value <0.05)

data_for_plot <- left_join(all_passed_data, data)

ggplot(data_for_plot, aes(Treatment, count)) +
  geom_boxplot(aes(color = gene_name)) +
  stat_compare_means(method = "t.test") +
  facet_wrap(~gene_name, scales = "free") +
  ggtitle("Significantly Different Genes (Normal Data)") +
  ylab("Gene Expression") +
  theme(legend.position = "none")
```

# Compare Saline between Timepoints

Normality test: these genes are the ones that are normally distributed
```{r}

saline_timepoint_data <- data %>%
  filter(Treatment == "Saline")

LPS_timepoint_data <- data %>%
  filter(Treatment == "LPS")

normal_test_data <- LPS_timepoint_data %>%
  group_by(gene_name) %>% 
  do(tidy(shapiro.test(.$count))) 

normally_distributed <- normal_test_data %>%
  filter(p.value > 0.05)

not_normal_distribution <- normal_test_data %>%
  filter(p.value < 0.05) %>%
  select(gene_name)

# run Mann-Whitney Wilcox on non-normally distributed
if (nrow(not_normal_distribution) > 0) {
data_for_Wilcox <- left_join(not_normal_distribution, LPS_timepoint_data) 

significant_Wilcox <- data_for_Wilcox %>%
  group_by(gene_name) %>% 
  do(tidy(wilcox.test(.$count~.$Time))) %>%
  filter(p.value < 0.0500)

wilcox_for_plot <- left_join(significant_Wilcox, LPS_timepoint_data)

ggplot(wilcox_for_plot, aes(Time, count)) +
  geom_boxplot(aes(color = gene_name)) +
  facet_wrap(~gene_name, scales = "free") +
  stat_compare_means(method = "wilcox.test", exact = TRUE) +
  ggtitle("ignificantly Different Genes Between Timepoints (Non-Normal Data)") +
  ylab("Gene Expression") +
  theme(legend.position = "none") }
```

F-test: tests if there are differences in the variances
```{r message = FALSE}
normally_distributed_remove <- normally_distributed %>%
  select(gene_name)

data_for_f_test <- left_join(normally_distributed_remove, LPS_timepoint_data)

f_test_pass <- data_for_f_test %>%
  group_by(gene_name) %>%
  nest() %>%
  mutate(f_test_result = map(data, ~var.test(count ~ Time, data = .x)),
         tidy_f_test = map(f_test_result, tidy)) %>%
  unnest(tidy_f_test, .drop = TRUE) %>%
  filter(p.value > 0.05)

```

T test: 
```{r}
f_test_genes <- f_test_pass %>%
  select(gene_name)

data_for_t_test <- left_join(f_test_genes, LPS_timepoint_data) 

t_test_data <- data_for_t_test %>%
  group_by(gene_name) %>% 
  do(tidy(t.test(.$count~.$Time)))

all_passed_data <- t_test_data %>%
  filter(p.value <0.05)

data_for_plot <- left_join(all_passed_data, LPS_timepoint_data)

ggplot(data_for_plot, aes(Time, count)) +
  geom_boxplot(aes(color = gene_name)) +
  stat_compare_means(method = "t.test") +
  facet_wrap(~gene_name, scales = "free") +
  ggtitle("Significantly Different Genes Between Timepoints (Normal Data)") +
  ylab("Gene Expression") +
  theme(legend.position = "none")
```
